---
title: "Reproducible Research Course Project 1"
author: "Lucas Gebhart"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/lucas.gebhart/Desktop/Coursera/reproducible project")
```

## This file conducts exploratory analysis of activity data.
Begin by reading in the data

```{r activity}
data <- read.csv("activity.csv")
head(data)
```

The data is every five minutes. Let's look at daily data
```{r}
dailysteps <- aggregate(steps~date, data, sum)
```

## Histogram of the steps data

Sometimes it helps to see a quick histogram.

```{r data, echo=FALSE}
hist(dailysteps$steps, main = "Histogram of Daily Steps", xlab="Daily Steps", ylab="Frequency")
```

The data appears pretty close to normally distributed, with a number of days where the person is inactive.
```{r}

mean(dailysteps$steps, na.rm= TRUE)
median(dailysteps$steps, na.rm= TRUE)

```
Clearly, a lot of data is missing and it is possible many of these 0s are incorrect. Let's look at the data on an interval basis, instead of a daily basis.

```{r}
meanintervalsteps <- aggregate(steps~interval, data, mean)
```

```{r}
plot(meanintervalsteps$interval, meanintervalsteps$steps, type = "l", main ="Average Step per Interval", xlab="5 minute interval", ylab="Mean Steps")
```

There is a big spike near 800-900 minutes into the day, let's find out what interval this is.

```{r}
moststeps <- max(meanintervalsteps$steps)
maxinterval <- data[meanintervalsteps$steps == moststeps,]
maxinterval[1,3]
```
The most steps occur around 835 minutes into a day. Maybe the person likes afternoon runs?
Now, let's deal with missing data... there are a few strategies of varying accuracy and ease of implementation. One option would be to take the average of the interval before and after the interval in question. I could easily implement this in excel... but have insufficent R skills.
The next option would be to give make the missing data the average. But there are clearly different activity patterns.
So let's fill the missing data with the average FOR THAT INTERVAL.

```{r}
missingdata <- is.na(data$steps)
head(missingdata)
data1<-data[is.na(data$steps),]
head(data1)

impute_int_avg <- meanintervalsteps$steps[match(data$interval, meanintervalsteps$interval)]
steps_replace <- transform(data, steps = ifelse(is.na(data$steps), yes = impute_int_avg, no = data$steps))

dailysteps_replace <- aggregate(steps~date, steps_replace, sum)
```
Let's see what the data looks like after replacing all of the missing values.

```{r}
hist(dailysteps_replace$steps, main = "Histogram of Daily Steps with imputed steps", xlab="Daily Steps", ylab="Frequency")
```
Let's see if there is a difference in activity between weekends and weekdays. We'll use data without imputed steps.

```{r}
#create the identifying column
data$weekday <- weekdays(as.Date(data$date))
data$weekend <- ifelse(data$weekday == "Saturday" | data$weekday == "Sunday", "Weekend", "Weekday")
head(data)

typemeanintervalsteps <- aggregate(steps~interval + weekend, data, mean)
View(typemeanintervalsteps)
weekendsteps <- subset(typemeanintervalsteps, weekend == "Weekend")
weekdaysteps <- subset(typemeanintervalsteps, weekend == "Weekday")



library(ggplot2)
plot2 <- ggplot()+
  geom_line(aes(x= typemeanintervalsteps$interval, y= typemeanintervalsteps$steps, group = typemeanintervalsteps$weekend, color = typemeanintervalsteps$weekend))+
  labs( x= "Interval",
        y= "Steps",
        color= "Legend")
plot2

```



